### 2015-02-10 ms/ms.Rmd
Ecology requires separate docs for authors and metadata, and in word

---
output:
  html_document
---

# *Ecological Archives* Exxxxx
# Falster D.S. *et al* BAAD: a Biomass And Allometry Database for woody plants. *Ecology* xxxx

@@

bib <- read.bibtex("refs.bib")
data <- baad$data
var_def <- baad$dictionary
vars <- setdiff(var_def$variable[var_def$type == "numeric"],
                c("map", "mat", "lai", "age"))
```

```{r,results="asis",echo=FALSE, message=FALSE}
writeLines(paste("**List of Authors:** ",
                 paste0(author_info$authors$name,
                        "^", author_info$authors$address_code, "^",
                        collapse = ", ")))
writeLines("\n\n")
writeLines(paste0(author_info$address_table$code, ". ",
                  author_info$address_table$address))
```

```{r precite, results='hide', message=FALSE, echo=FALSE}
# Kluge to avoid all author names being cited. Bug in knitcitations, see:
# https://github.com/cboettig/knitcitations/issues/65
citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))
citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))
citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))
citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a", "10.1016/j.tree.2013.05.002"))
citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.12688/f1000research.4518","10.1016/j.tree.2011.11.016"))
citep("10.1038/nj7461-243a")
citep(list(bib["Ishihara_Efficacy_2015"]))
```

## INTRODUCTION

Quantifying the amount of mass or energy invested in plant tissues is of fundamental interest across a range of disciplines, including ecology, forestry, ecosystem science, and climate change science `r citep(list(bib["niklas_plant_1994"],bib["chave_tree_2005"], "10.1111/j.1365-2745.2010.01735.x"))`. The allocation of net primary production into different plant components is an important process affecting the lifetime of carbon in ecosystems, and resource use and productivity by plants  `r citep(list(bib["cannell_carbon_1994"],bib["litton_carbon_2007"],bib["poorter_biomass_2012"]))`. While many scientific studies have destructively harvested woody plants, most of the data collected have only been made available in the form of summary tables or figures included in publications. Until now, the raw data has resided piecemeal on the hard drives of individual scientists spread around the world. Several studies have gathered together the fitted (allometric) equations for separate datasets `r citep(list("10.1016/S0378-1127(97)00019-4",bib["jenkins_national-scale_2003"],bib["zianis_biomass_2005"],"10.3832/ifor0901-006"))`, but none have previously attempted to organize and share the raw individual plant data underpinning these equations on a large scale. Gathered together, such data would represent an important resource for the community, meeting a widely recognized need for rich, open data resources to solve ecological problems `r citep(c("10.1016/j.tree.2013.05.002","10.1007/s13595-014-0375-3","10.1038/505160a"))`.

We (D.S. Falster and R.A. Duursma, with the help of D.R. Barneche, R.G. FitzJohn and A. Vårhammar) set out to create such a resource. We identified data sources from existing collaborations as well as a wide-ranging literature search, focussed on woody plants in field as well as controlled conditions. We constrained the search by including only data where a wide range in plant sizes was measured, as well as a set of minimum variables (plant height and leaf mass or area), and an emphasis on field-grown plants (nearly all data from plants grown in controlled conditions arose from existing collaborations). We do not claim to have done an exhaustive search of the literature, due to the very large number of potential studies that could be included. In each case we asked authors directly whether they would be willing to make their raw data files freely available. The response was overwhelming: nearly everyone we contacted was interested to contribute their raw data. Moreover, we were invited to incorporate another compilation led by M.I. Ishihara and focusing on Japanese literature `r citep(list(bib["Ishihara_Efficacy_2015"]))`. As a result, we present BAAD: a Biomass And Allometry Database for woody plants, comprising data collected in `r length(unique(data$studyName))` different published and unpublished studies (Fig. 1).

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
# World plot colored by vegetation type
cols <- make_transparent(c(
    "blue","brown","forestgreen","darkorange","dodgerblue2",
    "navajowhite4","lightgoldenrod3","lavender"),0.67)
vegs <- c("BorF","Sav","TempF","TempRF","TropRF","TropSF","Wo")
vegs_labels <- c("Boreal forest","Savannah","Temperate forest",
                 "Temperate rainforest","Tropical rainforest",
                 "Tropical seasonal forest","Woodland","Glasshouse / Common garden")

drawWorldPlot(data[data$vegetation == vegs[1],], sizebyn=TRUE, pchcol=cols[1])
for(i in 2:length(vegs)){
  drawWorldPlot(data[data$vegetation == vegs[i],], sizebyn=TRUE, pchcol=cols[i], add=TRUE)
}
drawWorldPlot(data[data$growingCondition == "GH",], sizebyn=TRUE,
              pchcol=cols[8], add=TRUE)
par(xpd=NA)
legend(-160, -100, vegs_labels[1:4], fill=cols[1:4], bty='n')
legend(-40, -100, vegs_labels[5:8], fill=cols[5:8], bty='n')
```

**Figure 1** Global distribution of studies included in the BAAD. Symbols are sized relative to the number of individual plants in each dataset.

Combining data from many different studies, each using its own set of terms and units, and quirks, is no trivial task `r citep(c("10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2","10.1016/j.tree.2011.11.016","10.12688/f1000research.4518"))`. This compilation would not have been possible without several new technologies. In particular,

1. Our entire work flow is scripted, using the R computing environment. The database can be rebuilt in an instant from the raw data files, using a single command. When a single data contribution has been modified, the final master dataset is thus easily recreated.
2. We minimised the amount of code by requiring each study to conform to a common format, with separate files for raw data, units of data, meta data, contact details, and citation (see examples in attached code). These files were then processed in a standardised way (Fig. 2).
3. We used the package `knitr` within `R` to automatically produce reports comparing the raw data provided by a single study to the rest of the database `r citep(citation("knitr"))`. These reports were sent to data contributors for verification, and thus allowed for easy identification of errors in the raw data or processing. Each report included plots for every pairwise combination of quantitative variables in a given study, site maps, and metadata. The updated reports are available on the database website.
4. We used version control and the code sharing website [github](https://github.com/dfalster/baad/) to track the evolution of the database and the code to process the raw data. Datasets as well as processing methods inevitably contain errors. With version control, we are able to trace these errors to the source, and review review the history of changes to any given dataset.

![baadworkflow](Figure2.png)
**Figure2** Workflow for building the BAAD. Data from each study is processed in the same way, using a standardised set of input files, resulting in a single dataset with a common format.

Until recently, most scientific publications in biology have mostly been concerned with "small data" `r citep("10.1038/nj7461-243a")`. Datasets arrive in small parcels, corresponding to the work of an individual study, or perhaps a handful of studies. It is exciting to be able to put these parcels together, to create something larger. We hope others will use this dataset and build on it. See [https://github.com/dfalster/baad](https://github.com/dfalster/baad) for information on how to contribute to future releases.

## METHODS

## METADATA CLASS I. DATA SET DESCRIPTORS

### A. DATA SET IDENTITY
the BAAD: a Biomass And Allometry Database for woody plants

### B. DATA SET IDENTIFICATION CODE:

### C. DATA SET DESCRIPTORS:
**1. Originators**: The study was initiated by D.S. Falster and R.A. Duursma. D.R. Barneche, R.G. FitzJohn and A. Vårhammar assisted with literature searching, data compilation, and coding. A separate compilation focusing on natural forests in Japan, was compiled by M.I. Ishihara, with assistance from T. Hiura, H. Utsugi and H. Tanouchi. All other authors contributed data to the project.

**2. Abstract**: Understanding how plants are plant constructed -- i.e. how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping energy-fluxes in the terrestrial biosphere. Allocation patterns can differ through ontogeny, but also among coexisting species, and among species adapted to different environments. While a variety of models dealing with biomass allocation exist, we lack a synthetic understanding of the underlying processes. This is partly due to the lack of suitable datasets for validating and parameterising models. To that end, we present the Biomass And Allometry Database (BAAD) for woody plants. The BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species. Most of these data come from an existing publication, however, raw data were rarely made public at time of publication. Thus the BAAD contains individual level data from different studies, transformed into standard units and variable names. The transformations were achieved using a common workflow for all raw data files. Other features which distinguish the BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e. not indirectly via allometric equations (except in very large trees where biomass was estimated from detailed sub-sampling). We included both wild and artificially grown plants. The dataset contains the following size metrics: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; height of plant and crown base, crown area and surface area; and the dry mass of leaf, stem, branches, sapwood, heartwood, bark, coarse roots and fine root tissues. We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, light, experimental treatment, vegetation type) where available. It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the world's vegetation.

### D. Key words

plant traits; biomass allocation; biomass partitioning; plant allometry; allometric equations; global carbon cycle

## METADATA CLASS II. RESEARCH ORIGIN DESCRIPTORS

### B. SPECIFIC SUBPROJECT DESCRIPTION

```{r,results="asis",echo=FALSE, message=FALSE}
details2B <- sapply(split(data, data$studyName), study_details, baad)
writeLines(details2B, sep="\n\n")
```

## CLASS III. DATA SET STATUS AND ACCESSIBILITY

### A. Status

**Latest update**: The dataset made available through this publication was processed on 2014-12-15. Newer updates may be retrieved from the BAAD website ([https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).

**Metadata status**: Metadata are complete to last update and are stored with data.

**Data verification**: The raw data passed through several rounds of review, to ensure units were correct, raw data was appropriately matched to variables in the BAAD, and to minimize outliers.

i) Initial analysis by data contributors. The majority of studies have been published and as such have gone through quality assurance by the data contributor.
ii) Review by compilers of database. We have gone through extensive data verification by checking cross-correlations, credible ranges, and outliers of all variables in the database.
iii) All species names were matched against online databases with the R packages `Taxonstand` and `taxize` R package `r citep(list("10.1111/j.2041-210X.2012.00232.x", bib["chamberlain2013taxize"]))`.
iv) Additional review by contributors using reports generated from the data.  In these reports, we summarized species, locations (latitude, longitude and maps), plant functional type, metadata provided by the contributor, and pair-wise plots of all numeric variables in the dataset (and a comparison to all data in the database). Each contributor checked the reports for problems, which allowed us to improve data quality in many cases.
v) Derived variables were calculated from the raw data where appropriate. For example, a dataset may contain crown length and plant height but not height to crown base, in which case the latter was calculated from the other two variables.

### B. Accessibility

**1. Storage location and medium**: The data published here is the first release from the Biomass and allometry database stored at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/). The version of the data made available for this publication is available here ((LINK TO ECOLOGICAL ARCHIVES DATA FILE)). For scientific analyses, we suggest you use this version.

Development versions of the dataset will continue to be available at [https://github.com/dfalster/baad/](https://github.com/dfalster/baad/).

If you notice any problems with the dataset, please list an issue [here](https://github.com/dfalster/baad/issues/new).

**Contact persons**: Queries about individual specific data points can be directed to the contributing author for that study. For queries about the entire dataset, please contact either Daniel Falster or Remko Duursma. For problems with the dataset, please submit an issue as described above.

**Copyright restrictions**: The dataset is released under the [Creative Commons CC0 license](http://creativecommons.org/choose/zero/?lang=en). As such, it may be freely used and redistributed. When using the dataset, we kindly request that you cite this article, recognizing the hard work that went into collecting the data and the authors' willingness to make it publicly available.

**Costs**: None.

## CLASS IV. DATA STRUCTURAL DESCRIPTORS

### A. Data Set Files

The BAAD contains the following elements:

- `data`: amalgamated dataset, with columns as defined in `dictionary`
- `dictionary`: a table of variable definitions
- `metadata`: a table with columns "studyName","Topic","Description", containing written information about the methods used to collect the data
- `methods`: a table with columns as in data, but containing a code for the methods used to collect the data. See [config/methodsDefinitions.csv](config/methodsDefinitions.csv) for codes.
- `references`:  as both summary table (giving DOI, URL, citation) and bibtex entries containing the primary source for each study
- `contacts`: table with contacts and affiliations for each study.

These elements are available as both

1) a list within the file `baad.rds` (for use in R)
2) as a series of csv and text files.

### B. Variable definitions for data.csv

```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
writeLines(variable_details(var_def))
```

**Number of observations by variable**

```{r results='asis', echo=FALSE, warning=FALSE}
options(digits=9)
kable(summary_table(data, var_def), "markdown", align="l", padding=10)
```

## ACKNOWLEDGMENTS

**Compilers:** D.S. Falster, A. Vårhammer and D.R. Barneche were employed on an ARC discovery grant to Falster (DP110102086) and a UWS start-up grant to R.A. Duursma. R.G. FitzJohn was supported by the Science and Industry Endowment Fund (RP04-174). M.I. Ishihara was supported by the Environmental Research and Technology Development Fund (S-9-3) of the Ministry of the Environment, Japan. We thank Takahiro Inoue, Hayato Aoyama and Shin-ichiro Aiba for valuable assistance.

In addition, we thank the following people for contributions to the following studies:
```{r,results="asis",echo=FALSE, message=FALSE}
writeLines(get_acknowledgements(baad))
```
.

## REFERENCES

```{r, echo=FALSE, results='asis'}
bibliography(sorting = 'ynt')
```

>>>>

# BAAD: a Biomass And Allometry Database for woody plants

@@

data <- baad$data
var_def <- baad$dictionary
vars <- setdiff(var_def$variable[var_def$type == "numeric"],
                c("map", "mat", "lai", "age"))
```

```{r,results="asis",echo=FALSE, message=FALSE}
writeLines(paste0(author_info$authors$name,
                        "^", author_info$authors$address_code, "^",
                        collapse = ", "))
writeLines("\n\n")
writeLines(paste0("^", author_info$authors$address_code, "^ ",
                  author_info$address_table$address,"  "))
```

Abstract. Understanding how plants are plant constructed -- i.e. how key size dimensions and the amount of mass invested in different tissues varies among individuals -- is essential for modeling plant growth, estimating carbon stocks, and mapping energy-fluxes in the terrestrial biosphere. Allocation patterns can differ through ontogeny, but also among coexisting species, and among species adapted to different environments. While a variety of models dealing with biomass allocation exist, we lack a synthetic understanding of the underlying processes. This is partly due to the lack of suitable datasets for validating and parameterising models. To that end, we present the Biomass And Allometry Database (BAAD) for woody plants. The BAAD contains `r sum(!is.na(data[,vars]))` measurements collected in `r length(unique(data$studyName))` different studies, from `r nrow(data)` individuals across `r length(unique(data$species))` species. Most of these data come from an existing publication, however, raw data were rarely made public at time of publication. Thus the BAAD contains individual level data from different studies, transformed into standard units and variable names. The transformations were achieved using a common workflow for all raw data files. Other features which distinguish the BAAD are: i) measurements were for individual plants rather than stand averages; ii) individuals spanning a range of sizes were measured; iii) inclusion of plants from 0.01-100m in height; and iii) biomass was estimated directly, i.e. not indirectly via allometric equations (except in very large trees where biomass was estimated from detailed sub-sampling). We included both wild and artificially grown plants. The dataset contains the following size metrics: total leaf area; area of stem cross-section, including sapwood, heartwood, and bark; height of plant and crown base, crown area and surface area; and the dry mass of leaf, stem, branches, sapwood, heartwood, bark, coarse roots and fine root tissues. We also report other properties of individuals (age, leaf size, leaf mass per area, wood density, nitrogen content of leaves and wood), as well as information about the growing environment (location, light, experimental treatment, vegetation type) where available. It is our hope that making these data available will improve our ability to understand plant growth, ecosystem dynamics, and carbon cycling in the world's vegetation.

Key words: plant traits; biomass allocation; biomass partitioning; plant allometry; allometric equations; global carbon cycle
